---
title: Assignment 2 - Setting up the computing environment. Due September 20, 2018.

# Use letters for affiliations
author:
  - name: EPIB607 - Inferential Statistics
    affiliation: a
#  - name: Second Author
#    affiliation: a,b
address:
  - code: a
    address: Fall 2018, McGill University
#  - code: b
#    address: Department of Neat Tricks, Whereever State University, Someplace, MC, 67890

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Author and Author

# Place DOI URL or CRAN Package URL here
doi: "https://sahirbhatnagar.com/EPIB607/"

# Abstract
abstract: |
  Computing is an essential part of modern statistics. However, before doing any data analysis, we must first install the necessary tools. In this assignment you will first signup for a free DataCamp account and login to your McGill GitHub account. Then you will complete a series of DataCamp courses that will guide you through installing R, RStudio and Git. After installation, you will learn how to use the RStudio IDE, create projects and run some basic commands. You will then be introduced to R Markdown which is a tool for creating reproducible reports. All future assignments for this course must be in R Markdown format and submitted to a private online GitHub repository. 

# Optional: Acknowledgements
#acknowledgements: |
#  This template package builds upon, and extends, the work of the excellent
#  [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
#  [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
#  gratefully acknowledged as this work would not have been possible without them.  Our extensions
#  are under the same respective licensing term
#  ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
#  [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).

# Optional: One or more keywords
keywords:
  - R
  - RStudio
  - Git
  - GitHub
  - R Markdown
  - DataCamp

# Paper size for the document, values of letterpaper and a4paper
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: pinp

# Optional: Enable a 'Draft' watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
footer_contents: "Assignment 1 due Sepetember 20, 2018 by 5pm"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{Due Sepetember 20, 2018}
  %\VignetteEngine{knitr::rmarkdown}
---

# Marking {-}

Your progress and completion of these courses will be availble to us automatically through the DataCamp website. **You do not need to hand anything in for this Assignment**. You will receive full credits for this assignment once we have seen that all tasks have been completed.

# Sign up for DataCamp and GitHub 

**Please try to complete this step before class on Thursday September 6**:  

1. Sign up for a free DataCamp account at [this link](https://www.datacamp.com/groups/shared_links/4c7d78a632b557dfdd6618b3e8fac09495571fec). Note: you are required to sign up with a `@mail.mcgill.ca` or `@mcgill.ca` email address.  
2. Sign in to [https://github.mcgill.ca/](https://github.mcgill.ca/) using your McGill email address and corresponding password.  




# Install Git

You need to first install the [git](https://git-scm.com/) version control system on your system. Follow Chapter 1: Installing Git [at this link](https://plot.ly/r/github-getting-started-for-data-scientists/#chapter-1-installing-git) for step-by-step installation instructions with screenshots. 




# Install R and RStudio

This short course will guide you through installing both [R](https://cran.r-project.org/) and [RStudio](https://www.rstudio.com/products/rstudio/download/preview/). RStudio is a software application that facilitates how you interact with `R`. Click on the link the in the figure caption join the course.

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=1in, height=1in]{../../images/rstudio_ide.png} 
  \end{center}
  \caption{\href{https://www.datacamp.com/courses/working-with-the-rstudio-ide-part-1}{Working with the RStudio IDE (Part 1)}}\label{fig}
\end{figure}


# Introduction to R

In this course you will get a hands-on introduction to the basic commands in R. With the knowledge gained in this course, you will be ready to perform a data analysis. Click on the link the in the figure caption join the course.  

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=1in, height=1in]{../../images/intro_r.png} 
  \end{center}
  \caption{\href{https://www.datacamp.com/courses/free-introduction-to-r}{Introduction to R}}\label{fig}
\end{figure}


# R Markdown

You will learn how to create reproducible reports using R and Markdown. All assignments for this course must be submitted in this format. Click on the link the in the figure caption join the course.  

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=1in, height=1in]{../../images/rmarkdown_r.png} 
  \end{center}
  \caption{\href{https://www.datacamp.com/courses/reporting-with-r-markdown}{Reporting with R Markdown}}\label{fig}
\end{figure}


# GitHub using RStudio

You will learn how to use RStudio to version control your code. All assignments for this course must be submitted to a GitHub repository. Click on the link the in the figure caption join the course (Chapter 2 only).

\begin{figure}[H]
  \begin{center}
    \includegraphics[width=1in, height=1in]{../../images/rstudio_ide_2.png} 
  \end{center}
  \caption{\href{https://www.datacamp.com/courses/working-with-the-rstudio-ide-part-2}{Version Control with RStudio IDE (Chapter 2 only)}}\label{fig}
\end{figure}



<!-- 
We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermarking;



## References 

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `jss.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code 

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures 

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.66\textwidth, height=3.5in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) 

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by three backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typeset directly simply by relying on pandoc.


## Single column equations 

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

