--- 
title: "MATH 697"
author: "Sahir Rai Bhatnagar"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: sahirbhatnagar/MATH697
description: "Course material for MATH 697"
---

# Syllabus {-}


## General Information {-}

- Instructor: Sahir Bhatnagar
- Email: [sahir.bhatnagar@mail.mcgill.ca](mailto:sahir.bhatnagar@mail.mcgill.ca)
- Website: [http://sahirbhatnagar.com/MATH697/](http://sahirbhatnagar.com/MATH697/)
- Lectures: Tuesdays 9am - 12pm
- Office: TBD
- Office Hours: By appointment only
- Prerequisite(s): Calculus and Algebra
- Texts: _Modern Mathematical Statistics with Applications_, 2nd Edition by Jay L. Devore and Kenneth N. Berk

## Course Description {-}

The main learning outcomes of this course are to get a broad idea about some frequently used probability models and to learn basic results and techniques in probability theory and statistical inference. Most of the materials for the course will be drawn from the first seven chapters of the textbook. The book does not, however, contain all the materials we intend to cover in this course. Some extra notes will therefore be given on those topics not in the text book. We will also introduce computational methods in statistics with the statistical software program `R`.



## Grade Distribution  {-}

```{r, results='asis', echo=FALSE}
knitr::kable(data.frame(c("Assignments","Quizzes","Final Exam"),
                        c("10%","40%","50%")), col.names = c("",""))
```

## Target Syllabus {-}

### Overview and Descriptive Statistics (Weeks 1-4)  {-}
- 1.1 Populations and Samples 
- 1.2 Pictorial and Tabular Methods in Descriptive Statistics 
- 1.3 Measures of Location 
- 1.4 Measures of Variability

### Probability (Weeks 1-4)  {-}
- 2.1 Sample Spaces and Events
- 2.2 Axioms, Interpretations, and Properties of Probability
- 2.3 Counting Techniques 
- 2.4 Conditional Probability 
- 2.5 Independence

### Discrete Random Variables and Probability Distributions (Weeks 1-4)  {-}
  
- 3.1 Random Variables 
- 3.2 Probability Distributions for Discrete Random Variables 
- 3.3 Expected Values of Discrete Random Variables 
- 3.4 Moments and Moment Generating Functions 
- 3.5 The Binomial Probability Distribution 
- 3.7 The Poisson Probability Distribution 

### Continuous Random Variables and Probability Distributions (Weeks 5-8) {-}

- 4.1 Probability Density Functions and Cumulative Distribution Functions 
- 4.2 Expected Values and Moment Generating Functions 
- 4.3 The Normal Distribution 
- 4.7 Transformations of a Random Variable 


### Joint Probability Distributions (Weeks 5-8) {-}

- 5.1 Jointly Distributed Random Variables 
- 5.2 Expected Values, Covariance, and Correlation 
- 5.3 Conditional Distributions 
- 5.4 Transformations of Random Variables 

### Statistics and Sampling Distributions (Weeks 5-8) {-}

- 6.1 Statistics and Their Distributions 
- 6.2 The Distribution of the Sample Mean 
- 6.3 The Mean, Variance, and MGF for Several Variables 
- 6.4 Distributions Based on a Normal Random Sample 

### Point Estimation (Weeks 9-12) {-}

- 7.1 General Concepts and Criteria 
- 7.2 Methods of Point Estimation 

### Statistical Intervals Based on a Single Sample (Weeks 9-12) {-}
- 8.1 Basic Properties of Confidence Intervals 
- 8.2 Large-Sample Confidence Intervals for a Population Mean and Proportion 
- 8.3 Intervals Based on a Normal Population Distribution 
- 8.4 Confidence Intervals for the Variance and Standard Deviation of a Normal Population 
- 8.5 Bootstrap Confidence Intervals 

### Tests of Hypotheses Based on a Single Sample (Weeks 9-12) {-}

- 9.1 Hypotheses and Test Procedures 
- 9.2 Tests About a Population Mean 
- 9.3 Tests Concerning a Population Proportion 
- 9.4 P-Values 
- 9.5 Some Comments on Selecting a Test Procedure 



# Prerequisites {-}

## Install R and RStudio  {-}

All examples in this book are run in an [R](https://cran.r-project.org/) environment. You also need a recent version of [RStudio](https://www.rstudio.com/products/rstudio/download/preview/), which is a software application that facilitates how you interact with `R`. It is developed by data enthusiasts who consider statistics to be more than just simulations, formulas and proofs. RStudio emphasizes the following: 

1. **Version control**:  [Why I should use version control](http://stackoverflow.com/questions/1408450/why-should-i-use-version-control) especially for the [solo data analyst](http://stackoverflow.com/questions/2712421/r-and-version-control-for-the-solo-data-analyst).

2. **Reproducible research**: seamless integration with [RMarkdown](http://rmarkdown.rstudio.com/) for creating [dynamic documents](http://yihui.name/knitr/) and presentations

3. **Creating R Packages**: seamless integration with the [devtools](https://github.com/hadley/devtools) package for creating software that implements your statistical method or analysis. 



## R Packages  {-}

The following packages will be called upon at some point, so please install them before getting started with the tutorials. Enter the following command in `R`:

```{r eval=FALSE, echo=TRUE}
install.packages(c("pacman","knitr","data.table", "rmarkdown", "tidyverse", "boot", "Hmisc"))
```


## Introduction to R {-}

Try out the interactive tutorial: [http://swirlstats.com/](http://swirlstats.com/)


## Background Reading  {-}

The greatest thing about `R` is that there are so many people out there willing to help you. `R` users are constantly writing tutorials and creating packages to make your analysis tasks easier. Here is a very targeted list that I suggest reading prior to starting the tutorials

1. [Writing Functions](http://r4ds.had.co.nz/functions.html) 
2. [`for` loops](http://r4ds.had.co.nz/iteration.html)
3. [`apply` vs. `for`](https://kbroman.wordpress.com/2013/04/02/apply-vs-for/)



```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'tidyverse', 'data.table', 'Hmisc', 'boot'), 
  'packages.bib')
```
