---
title: DALITE Q1 - Histograms, Medians and Means. Due September 12, 2018.

# Use letters for affiliations
author:
  - name: EPIB607 - Inferential Statistics
    affiliation: a
#  - name: Second Author
#    affiliation: a,b
address:
  - code: a
    address: Fall 2018, McGill University
#  - code: b
#    address: Department of Neat Tricks, Whereever State University, Someplace, MC, 67890

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Bhatnagar and Hanley

# Place DOI URL or CRAN Package URL here
doi: "https://sahirbhatnagar.com/EPIB607/"

# Abstract
abstract: |
  This DALITE quiz will cover the basic concepts of histograms, means and medians.  

# Optional: Acknowledgements
#acknowledgements: |
#  This template package builds upon, and extends, the work of the excellent
#  [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
#  [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
#  gratefully acknowledged as this work would not have been possible without them.  Our extensions
#  are under the same respective licensing term
#  ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
#  [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).

# Optional: One or more keywords
keywords:
  - Histogram
  - Density Plot
  - Mean
  - Median
  - Mode

# Paper size for the document, values of letterpaper and a4paper
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
#one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: pinp

# Optional: Enable a 'Draft' watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
footer_contents: "DALITE Q1 due Sepetember 12, 2018 by 5pm"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{Due Sepetember 12, 2018}
  %\VignetteEngine{knitr::rmarkdown}
---

# Marking {-}

Completion of this DALITE exercise will be availble to us automatically through the DALITE website. Therefore **you do not need to hand anything in**. Marks will be based on the number of correct answers. For each question you will receive 0.5 marks for getting the correct answer on the first attempt and an additional 0.5 marks if you stick with the right answer or switch to the correct answer after seeing someone else's rationale. 

# Sign up for DALITE 

**This step only needs to be completed once for the whole semester**:  

1. You can join the EPIB607 group by accessing the unique link: [https://mydalite.org/en/live/signup/form/NTc4](https://mydalite.org/en/live/signup/form/NTc4)
2. Upon accessing the link, you will be prompted to enter an e-mail address. I recommend using the same e-mail address as your DataCamp account. 
3. You never need to remember a username or password to access the DALITE platform; access to your assignments is managed through tokens sent to your e-mail address. You will be sent a new link everytime a new exercise has been posted. 

# Histograms

## Learning Objectives
1. Understand that the distribution of a variable consists of what values the variable takes and
how often.
2. Understand that class intervals should be of equal width; choose appropriate class widths
to effectively reveal informative patterns in the data.
3. Understand that the vertical axis of the histogram may be scaled for frequency, proportion,
or percentage. The choice of vertical scaling for any data set does not affect the important
features revealed by a histogram.
4. Be able to describe a graphical display of data by first describing the overall pattern and
then deviations from that pattern. Describe the shape of the overall pattern and identify any
gaps in data and potential outliers.
5. Recognize rough symmetry and clear skewness in the overall pattern of a distribution

## Videos

1. [Against All Odds Unit 3](https://www.learner.org/courses/againstallodds/unitpages/unit03.html)  

\vspace*{0.25cm}

## Required Readings

1. [Against All Odds Unit 3, pages 1-6](https://www.learner.org/courses/againstallodds/pdfs/AgainstAllOdds_StudentGuide_Unit03.pdf#page=1)
2. [Visualizing distributions: Histograms and density plots](https://serialmentor.com/dataviz/histograms-density-plots.html)


# Mean and Median

## Learning Objectives

1. Understand that graphical descriptions of data are more meaningful when supplemented
with numerical measures of center.
2. Know that the median (midpoint or typical value) and mean (arithmetic average) are
common measures of center (or location) for a distribution. 
3. Understand the formulas used to calculate the median, mean, and mode.
4. Know that the mean and median should be close in symmetric distributions and that the
mean is pulled toward the long tail of a skewed distribution. Know that the mean is a nonresistant measure of center because it is strongly influenced by extreme observations and that
the median is a resistant measure of center.
5. Be able to choose an appropriate measure of center in practice. 

## Videos

1. [Against All Odds Unit 4](https://www.learner.org/courses/againstallodds/unitpages/unit04.html)


## Required Readings

\begin{enumerate}
\item \href{https://www.learner.org/courses/againstallodds/pdfs/AgainstAllOdds_StudentGuide_Unit04.pdf#page=1}{Against All Odds Unit 4, pages 1-6}
\end{enumerate}







<!-- 
We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermarking;



## References 

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `jss.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code 

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures 

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.66\textwidth, height=3.5in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) 

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by three backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typeset directly simply by relying on pandoc.


## Single column equations 

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

