---
title: DALITE Q5 - Bootstrap, Tests of Significance and Small Sample Inference for One Mean. Solutions.

# Use letters for affiliations
author:
  - name: EPIB607 - Inferential Statistics
    affiliation: a
#  - name: Second Author
#    affiliation: a,b
address:
  - code: a
    address: Fall 2018, McGill University
#  - code: b
#    address: Department of Neat Tricks, Whereever State University, Someplace, MC, 67890

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Bhatnagar and Hanley

# Place DOI URL or CRAN Package URL here
doi: "https://sahirbhatnagar.com/EPIB607/"

# Abstract
abstract: |
  This DALITE quiz will cover the bootstrap, an introduction to significance testing, and inference for a single mean using the t distribution. 

# Optional: Acknowledgements
#acknowledgements: |
#  This template package builds upon, and extends, the work of the excellent
#  [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
#  [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
#  gratefully acknowledged as this work would not have been possible without them.  Our extensions
#  are under the same respective licensing term
#  ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
#  [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).

# Optional: One or more keywords
keywords:
  - Hypothesis testing
  - Bootstrap
  - t distribution
  - One sample mean
  - Normal calculations
  - Confidence intervals
  - Central Limit Theorem (CLT)

# Paper size for the document, values of letterpaper and a4paper
papersize: letter

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: pinp

# Optional: Enable a 'Draft' watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
footer_contents: "DALITE Q5 due October 3, 2018 by 5pm"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{Due Sepetember 26, 2018}
  %\VignetteEngine{knitr::rmarkdown}
---

# Hypothesis tests 1

The average human gestation time is 266 days from conception. A researcher suspects that proper nutrition plays an important role and that poor women with inadequate food intake would have shorter gestation times even when given vitamin supplements. A random sample of 20 poor women given vitamin supplements throughout the pregnancy has a mean gestation time from conception of $\bar{y}$=256 days. The null hypothesis for the researcher's test is


a. **$\textrm{H}_0$: $\mu$ = 266 (Correct)**
b. $\textrm{H}_0$: $\mu$ = 256
c. $\textrm{H}_0$: $\mu$ < 266
d. $\textrm{H}_0$: $\bar{y}$ = 266


## Correct rationales

- The null hypothesis is usually a statement of "no effect" or "no difference."
- The null hypothesis is the statement of no effect, so if there was no effect, the average gestation time would be the population mean.
- The null hypothesis of this study is that the average gestation period of women with proper nutrition and poor women with inadequate food intake is the same that is 266 days. Hence, the population mean (all poor women) have the gestation period of 266 days.


## Incorrect rationales

- The null hypothesis is "the accepted fact", which in this case is that poorer nutrion will yield results lower than the mean (<266) or H null: u<266
- The researcher is testing that the hypothesis that poor women with inadequate food intake will have lower gestation time

# Hypothesis tests 2

The average human gestation time is 266 days from conception. A researcher suspects that proper nutrition plays an important role and that poor women with inadequate food intake would have shorter gestation times even when given vitamin supplements. A random sample of 20 poor women given vitamin supplements throughout the pregnancy has a mean gestation time from conception of $\bar{y}$=256 days. The researcher's alternative hypothesis for the test is

a. $\textrm{H}_a$: $\mu \neq 256$
b. **$\textrm{H}_a$: $\mu < 266$ (Correct)**
c. $\textrm{H}_a$: $\mu < 256$


## Correct rationales

- The alternative hypothesis is the point of view of the researchers. In this scenario, the researchers are suspecting that there will be a shorter gestation time for poor women in comparison to an average woman. The alternative hypothesis should be one-sided because the investigators only want to know if gestation time will be shorter and not if it is overall different (shorter or greater) from the average gestation time.
- Researcher has specified a specific direction - that the sample would generate a mean gestation time less than 266. Therefore the alternative hypothesis is that $\mu$ is less than 266.

## Incorrect rationales



# Hypothesis test 3

The average human gestation time is 266 days from conception. A researcher suspects that proper nutrition plays an important role and that poor women with inadequate food intake would have shorter gestation times even when given vitamin supplements. A random sample of 20 poor women given vitamin supplements throughout the pregnancy has a mean gestation time from conception of $\bar{y}$=256 days. Human gestation times are approximately Normal with standard deviation $\sigma$ = 16 days. The p-value for the researcher's test is (provide your calculation in the rationale)


a. more than 0.1
b. **less than 0.01 (Correct)**
c. less than 0.001
d. less than 0.05
e. less than 0.025




## Correct rationales

- $Z= (y - \mu)/ (\sigma / sqrt (n))$ = -2.795 $\to$ `pnorm(-2.795, mean = 0 , sd=1 ) = 0.00026`
- `mosaic::xpnorm(q= (256 - 266)/(16/sqrt(20)))`
- `stats::pnorm(q = 256, mean = 266, sd = 16/(sqrt(20)), lower.tail = T)`



## Incorrect rationales

- $t* = (256-266) / (16 / \sqrt(20))$ = -2.795. degrees of freedom = n-1 = 19. p-value is between 0.01 and 0.005. This is a one tailed t-test so divide by 2 and p-value is between 0.005 and 0.0025. Therefore, less than 0.01.
- $(256 - 266)/16 = -0.625 \to$ `pnorm(-0.625)` corresponds to a p value of 0.27
- $t = (\bar{y}-\mu)/(\sigma/\sqrt{n})$ then use $t$ distribution with df=19



# Hypothesis tests 4

Average human gestation time is 266 days, when counted from conception. A hospital gives a 90% confidence interval for the mean gestation time from conception among its patients. That interval is 264 $\pm$ 5 days. Is the mean gestation time in that hospital significantly different from 266 days?

a. **It is not significantly different at the 10% level and therefore is also not significantly different at the 5% level. (Correct)**
b. It is not significantly different at the 10% level but might be significantly different at the 5% level. c. It is significantly different at the 10% level 


## Correct rationales

- The average human gestation time falls within the interval of the hospital which is 259-269 days, therefore the mean gestation time is not significantly different at the 10% level and since the 5% level is even more specific, then it is not significantly different at the 5% level either.
- 266 is included in the 259-269 range, therefore the mean gestation time in the hospital is not significantly different from 266 days at the 10% level. Looking at the significance at the 5% level would make the confidence interval bigger, meaning there's not significant different there either

## Incorrect rationales

- Because 266 falls at the upper tail of the 90% confidence interval, therefore it is significant different at 10% significance level
- It is significantly different at the 10% level, since the CI does not contain the value zero.
- The 90% CI includes 266, but the 95% CI may not include 266, in which case a value like 266 would only be observed less than or equal to 5% of the time.
- To be significantly different, p-value should be less than 5%.
- As 266 falls within 264+/- 5 days, it is not significant at the 10% level, but it may not fall under this range because when the level of confidence is increased, the confidence interval becomes smaller.



# One sample mean

We prefer the $t$ procedures to the $z$ procedures for inference about a population mean because

a. $z$ can be used only for large samples
b. **$z$ requires that you know the population standard deviation $\sigma$ (Correct)**
c. $z$ requires that you can regard your data as an SRS from the population


## Correct rationales

- The reason we can use z on large samples is because when we have large samples, we can use the sample sd as an estimate for the population sd. The core reason for using t is because we don't know the population sd and we can't estimate the sd with the small sample size.
- because you don't know population sd, only sample sd for t procedures
- Z test needs to have a known population standard deviation sigma and either a normal distribution or the sample size n is large. If the population standard deviation is unknown and the sample size is small then t test statistic can be used.

## Incorrect rationales

- df are smaller for t-test, and you do not need a SD value


# One sample mean 2

Because $t$ procedures are robust, the most important condition for their safe use is that

a. the population standard deviation $\sigma$ is known
b. the population distribution is exactly Normal
c. **the data can be regarded as an SRS from the population (Correct)**
d. the CLT hasn't kicked in yet
e. the sample size is small 



## Correct rationales

- The point of the t test is that sigma is not known (rules out A) The CLT says that the population distribution doesn't matter (rules out B) We want the CLT to kick in (rules out D) The larger the sample size the better(rules out E).

## Incorrect rationales

- We should only use t procedures if the CLT is not applicable.
- In order to use the T procedure comfortably, the sample size must be large and the population distribution must be normal.

# Bootstrap

Which of the following statements _best_ describes the utility of the bootstrap

a. The bootstrap frees us from the requirement of using simple formulas to derive confidence intervals
b. The bootstrap allows us to simulate a sampling distribution
c. **The bootstrap frees us from the assumption of a Gaussian sampling distribution for the mean (as per the CLT) (Correct)**
d. The bootstrap tells us if the sampling distribution is asymmetric



## Correct rationales

- Computer intensive methods can solve most problems without assuming that the data have a Gaussian distribution.
- The bootstrap frees us from the assumption that the data conform to a bell-shaped normal distribution.

## Incorrect rationales

- due to the nature of bootstrap, you can resample things MANY times, which means that CLT will always kick in giving a gaussian distribution instead of assuming one exists
- Bootstrap is useful in because it does not require the CLT and **the population** does not have to be normally distributed
- The bootstrap doesn't rely on CLT, and gives us freedom from having to assume the normal distribution of **the population**.
- The bootstrap allows us to derive estimates when often-used theories do not apply. This is especially useful when n is small (e.g. CLT requires n > 30 to assume normality).

<!-- 
We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermarking;



## References 

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `jss.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code 

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures 

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.66\textwidth, height=3.5in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) 

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by three backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typeset directly simply by relying on pandoc.


## Single column equations 

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

