---
title: Assignment 8 - Comparing Two Means. Due November 16, 11:59pm 2018
author:
  - name: EPIB607 - Inferential Statistics
    affiliation: a
#  - name: Second Author
#    affiliation: a,b
address:
  - code: a
    address: Fall 2018, McGill University
#  - code: b
#    address: Department of Neat Tricks, Whereever State University, Someplace, MC, 67890
lead_author_surname: Bhatnagar and Hanley
doi: "https://sahirbhatnagar.com/EPIB607/"
abstract: |
  In this assignment you will practice comparing two means in a regression framework. Be sure to state the regression model in terms of parameters first. Use regression functions to fit these models. Answers should be given in full sentences (DO NOT just provide the number). All figures should have appropriately labeled axes, titles and captions (if necessary). Units for means and CIs should be provided. All graphs and calculations are to be completed in an R Markdown document. Please submit both the compiled HTML report and the source file (.Rmd) to myCourses by November 16, 2018, 11:59pm. Both HTML and .Rmd files should be saved as 'IDnumber_LastName_FirstName_EPIB607_A8'. 
# Optional: Acknowledgements
#acknowledgements: |
#  [rticles](https://cran.r-project.org/package=rticles) package, and both packages rely on the
#  [PNAS LaTeX](http://www.pnas.org/site/authors/latex.xhtml) macros. Both these sources are
#  ([GPL-3](https://www.gnu.org/licenses/gpl-3.0.en.html) and
#  [LPPL (>= 1.3)](https://www.latex-project.org/lppl/)).
# Optional: One or more keywords
keywords:
  - Two sample mean
  - Regression
  - Indicator variable
  - Bootstrap
papersize: letter
fontsize: 9pt
# Optional: Force one-column layout, default is two-column
one_column: true
# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true
# Optional: Enable one-sided layout, default is two-sided
#one_sided: true
# Optional: Enable section numbering, default is unnumbered
numbersections: true
# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5
# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true
# Optional: Bibliography 
bibliography: pinp
# Optional: Enable a 'Draft' watermark on the document
watermark: false
footer_contents: "Assignment 8 due November 16, 2018 by 11:59pm"
output: pinp::pinp
# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{Due November 16, 2018}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

# Template {-}

There is no template for this assignment. You may use the same template from previous assignments. 


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
```



# Food counseling for obese children

According to the Centers for Disease Control and Prevention (CDC), obesity rates among U.S. children have increased dramatically over the past three decades, from a low of about 5% to 18% as of 2010. A study examined the effect of family-based food-counseling sessions provided by trained professionals. The study randomly assigned obese children aged 9 to 12 years to either the counseling intervention or a control group not receiving any food counseling. The children's weight changes (in pounds, lb) after 15 weeks is displayed in Table \ref{fig:tab1}. 

```{r, echo=FALSE, eval=FALSE}
int <- c(-16.7,-14.8,-11.9,-9.7,-9.6,-8.8,-8.0,-7.1,-6.6,-6.0,-5.6,-5.6,
         -5.5,-5.5,-5.1,-5,-5,-4.8,-4.4,-4.4,-4.1,-4,-4,-3.6,-3.5,-3.2,-2.8,
         -2.0,-1.8,-1.8,-1.4,-1.2,-.2,-.1,0,0.2,0.6,1,1.2,1.2,1.4,1.8,2,2.2,
         2.5,2.8,3.3,4.2,5.4,5.8,6,6.4,8.4)
cont <- c(12,10,9,8.6,7.1,6.7,6.1,4.8,4.6,4.5,2.8,2.8,2.8,2.7,1.8,1.6,1,0.7,.4,.1,.1,-5.1,-5.1)
df <- data.frame(weight = c(int, cont), intervention = c(rep(1, length(int)),rep(0, length(cont))))

mean(int)
mean(cont)
fit <- glm(weight ~ intervention, data = df, family = gaussian(link="log"))
summary(fit)
```


a. Plot the data.
b. We are interested in knowing if obese children receiving food counseling gain less weight over a 15-week period. State the regression model as a function of population parameters. Be sure to define what each parameter represents. 
c. Use a regression procedure to answer this question. State the hypotheses, interpret the parameter of interest and a brief statement about your findings. 
d. Show how the \texttt{Std. Error} was calculated for the parameter of interest. 


\begin{figure}
  \begin{center}
    \includegraphics[scale=0.35, angle=90]{a83-crop.pdf} 
  \end{center}
  \caption{Weight change (lb) of obese children, by group}\label{fig:tab1}
\end{figure}

\clearpage

# Student drinking

A professor asked her sophomore students, "How many drinks do you typically have per session" (A drink is defined as one 12-ounce beer, one 4-ounce glass of wine, or one 1-ounce shot of liquor.) Some of the students didn't drink. Table \ref{fig:tab2} gives the responses of the female and male students who did drink. It is likely that some of the students exaggerated a bit. The sample is all students in one large sophomore-level class. The class is popular, so we are tentatively willing to regard its members as an SRS of sophomore students at this college.

a. Do a complete analysis that reports on a comparison of the drinking behavior of women and men.
b. Summarise your findings in a 140 character tweet.
c. (BONUS) Verify the constant variance assumption of your model.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.55, angle=90]{a81-crop.pdf} 
  \end{center}
  \caption{Drinks per session claimed by female and male students}\label{fig:tab2}
\end{figure}


\clearpage


# Ink toxicity

The National Toxicology Program evaluates the toxicity of chemicals found in manufacturing, in consumer products, or in the environment after disposal. Toxicity is assessed through a battery of tests. Here are some results from a study of the toxicity of black newsprint ink in 7-week-old female rats. The rats' fur was locally clipped twice a week for 13 weeks.
One group of rats received a dermal application of ink right after each clipping, and a control group of rats was left untreated. Table \ref{fig:tab3} shows the body weights (in grams) of female rats at the beginning of the study and at the end of the 13 weeks.

a. Verify that the two experimental groups are not significantly different at the beginning of the study
b. Is there good evidence that ink application impairs growth in female rats between 7 and 20 weeks of age? State your answer in terms of the \% difference between the mean weight gains of the two populations. Provide a 95% confidence interval for the \% difference between the mean weight gains of the two populations.
c. Show how the p-value in the \texttt{glm} output was calculated for the parameter of interest.


```{r,eval=FALSE, echo=FALSE}
int1 <- c(107.3,116.7,112.2,103.4,113.2,110.6,110.6,100.5,106.3,112.5)
int2 <- c(187,189.5,179.2,172.2,178.7,180.9,188.3,188.9,183.1,184.5)

cont1 <- c(111.2,105.4,110.8,105.6,106.1,104.4,114,115.1,109.2,111.3)
cont2 <- c(191.6,191.2,210.7,185.2,195,188.3,188.4,195.6,204.6,195.7)

df <- data.frame(weight = c(int2-int1, cont2-cont1), treat = c(rep(1, length(int1)), rep(0, length(cont1))))
df
fit <- glm(weight ~ treat, data = df, family = gaussian)
fit <- glm(weight ~ treat, data = df, family = gaussian(link = log))
summary(fit)
exp(-0.14370)
mean(int2-int1)/mean(cont2-cont1)
```


\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.55, angle=90]{a82-crop.pdf} 
  \end{center}
  \caption{Body weights (g) at beginning and end of ink toxicity study}\label{fig:tab3}
\end{figure}


<!-- 
We support several options via the YAML header

- Setting a DOI or URL footer, for example for the CRAN package URL,
  which is placed in the bottom-left footer of the title page and even pages;
- Setting a footer label, for example _YourPackage Vignette_ stating
  your package, which is placed in the bottom-right footer on odd
  pages;
- Setting a free-form author field used on the inside footer;
- Optional _Draft_ watermarking;



## References 

Here we differ from PNAS and suggest natbib. References will appear in
author-year form. Use `\citet{}`, `\citep{}`, etc as usual.

We default to the `jss.bst` style. To switch to a different bibliography
style, please use `biblio-style: style` in the YAML header.


## Inline R Code 

The PNAS sample included a fixed PNG image here, but this document prefers
to show the results and embedding of _R_ code. 

```{r figex, fig.width=3, fig.height=3, cache=TRUE, echo=TRUE, fig.cap="Narrow ggplot2 figure"}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) +
    geom_point(size=3, aes(colour=factor(cyl))) +
    theme(legend.position="none")
```

Here we use a standard knitr bloc with explicit options for

- figure width and height (`fig.width`, `fig.height`), both set to three inches;
- whether the code is shown (`echo=TRUE`); and
- the caption (`fig.cap`) as shown above.


## Digital Figures 

Markdown, Pandoc and LaTeX support `.eps` and `.pdf` files.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

The R examples above show how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. 

One (roundabout) way of doing this is to _not_ actually plot a figure, but to
save it in a file as the following segment shows:

```{r densityPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(data = midwest,
            mapping = aes(x = area,
                          fill = state,
                          color = state)) +
    geom_density(alpha = 0.3)
## save to file
suppressMessages(ggsave("densities.pdf", p))
```

This file is then included via standard LaTeX commands.

\begin{figure*}
  \begin{center}
    \includegraphics[width=0.66\textwidth, height=3.5in]{densities} 
  \end{center}
  \caption{Wide ggplot2 figure}\label{fig}
\end{figure*}


## Typeset Code (But Do Not Run It) 

We can also just show code.

```r
xx <- faithful[,"eruptions"]
fit <- density(xx)
plot(fit)
```

This simply used a pandoc bloc started and ended by three backticks,
with `r` as the language choice.  Similarly, _many_ other languages
can be typeset directly simply by relying on pandoc.


## Single column equations 

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \ref{eqn:example} below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

\begin{equation}
  \begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \\
       &=x^3+3x^2y+3xy^3+x^3. 
       \label{eqn:example} 
  \end{aligned}
\end{equation}


pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

