\documentclass[letterpaper,9pt,twoside,printwatermark=false]{pinp}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% The geometry package layout settings need to be set here...
\geometry{layoutsize={0.95588\paperwidth,0.98864\paperheight},%
          layouthoffset=0.02206\paperwidth,%
		  layoutvoffset=0.00568\paperheight}

\definecolor{pinpblue}{HTML}{185FAF}  % imagecolorpicker on blue for new R logo
\definecolor{pnasbluetext}{RGB}{101,0,0} %



\title{Assignment 9 - Poisson Regression. Due November 23, 11:59pm 2018}

\author[a]{EPIB607 - Inferential Statistics}

  \affil[a]{Fall 2018, McGill University}

\setcounter{secnumdepth}{5}

% Please give the surname of the lead author for the running footer
\leadauthor{Bhatnagar and Hanley}

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  Poisson regression |  Rate ratio |  Rate difference |  Person time |  Offset  }  

\begin{abstract}
In this assignment you will practice Poisson regression. Be sure to
state the regression model in terms of parameters first. Use regression
functions to fit these models. Answers should be given in full sentences
(DO NOT just provide the number). All figures should have appropriately
labeled axes, titles and captions (if necessary). Units for means and
CIs should be provided. All graphs and calculations are to be completed
in an R Markdown document. Please submit both the compiled HTML report
and the source file (.Rmd) to myCourses by November 23, 2018, 11:59pm.
Both HTML and .Rmd files should be saved as
`IDnumber\_LastName\_FirstName\_EPIB607\_A9'.
\end{abstract}

\dates{This version was compiled on \today}
\doi{\url{https://sahirbhatnagar.com/EPIB607/}}

\pinpfootercontents{Assignment 9 due November 23, 2018 by 11:59pm}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.


\section*{Template}\label{template}
\addcontentsline{toc}{section}{Template}

There is no template for this assignment. You may use the same template
from previous assignments. Be sure to include your code at the end of
the compiled report.

\section{Bednets}\label{bednets}

See the 2018 Lancet article
\textit{Efficacy of Olyset Duo, a bednet containing pyriproxyfen and permethrin, versus a permethrin-only net against clinical malaria in an area with highly pyrethroid-resistant vectors in rural Burkina Faso: a cluster-randomised controlled trial}
(\texttt{Bednets.pdf} in \texttt{A9} folder of myCourses) by Tiono et.
al.

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  Reproduce the Rate ratio and 95\% CI in Table 2. Start by writing down
  the model in terms of a regression equation with parameters and
  determinants. Interpret the rate ratio.
\item
  Perform a goodness of fit test for the model in part (a). Is this a
  good fit? If not, \underline{explain} what you might do to improve the
  fit.
\item
  Calculate the rate difference and 95\% CI comparing PPF-treated to
  Standard long-lasting insecticidal nets.
\end{enumerate}

\section{Population mortality rates in
Denmark}\label{population-mortality-rates-in-denmark}

The following table contains mortality data for males and females in
Denmark for 4 age groups over three time periods.

\vspace*{0.3in}

\begin{tabular}{l|l|r|r|r|r|r|r}
\hline
Year & Age & Female\_deaths & Female\_PT & Female\_rate & Male\_deaths & Male\_PT & Male\_rate\\
\hline
1980-1984 & 70-74 & 15989 & 586882.8 & 0.0272439 & 23810 & 456908.21 & 0.0521111\\
\hline
1980-1984 & 75-79 & 20838 & 454142.7 & 0.0458843 & 24707 & 300318.92 & 0.0822692\\
\hline
1980-1984 & 80-84 & 24073 & 297678.6 & 0.0808691 & 20319 & 167303.51 & 0.1214499\\
\hline
1980-1984 & 85-89 & 20216 & 147771.7 & 0.1368057 & 13524 & 74295.83 & 0.1820291\\
\hline
2000-2004 & 70-74 & 13912 & 521561.9 & 0.0266737 & 17360 & 436994.92 & 0.0397259\\
\hline
2000-2004 & 75-79 & 19731 & 471945.5 & 0.0418078 & 22477 & 341362.82 & 0.0658449\\
\hline
2000-2004 & 80-84 & 25541 & 369989.9 & 0.0690316 & 22992 & 217929.72 & 0.1055019\\
\hline
2000-2004 & 85-89 & 27135 & 226798.1 & 0.1196439 & 17444 & 104009.58 & 0.1677153\\
\hline
2005-2009 & 70-74 & 12179 & 540568.6 & 0.0225300 & 15782 & 472012.84 & 0.0334355\\
\hline
2005-2009 & 75-79 & 17273 & 444474.2 & 0.0388616 & 19547 & 344351.34 & 0.0567647\\
\hline
2005-2009 & 80-84 & 23513 & 363534.1 & 0.0646789 & 21781 & 230530.24 & 0.0944822\\
\hline
2005-2009 & 85-89 & 26842 & 237877.3 & 0.1128397 & 17811 & 114485.04 & 0.1555749\\
\hline
\end{tabular}

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  Come up with a suitable regression model for this data. Write down the
  regression equation in terms of parameters and determinants.
\item
  Estimate the parameters of this model using the data in the table
  above. Provide the fitted regression equation.
\item
  Interpret the parameter for gender. Are mortality rates significantly
  different in males compared with females?
\item
  Perform a goodness of fit test for the fitted model in part (b). Is
  this a good fit?
\end{enumerate}

%\showmatmethods


\bibliography{pinp}
\bibliographystyle{jss}



\end{document}

